Multiple Linear Regression on Sleep Data

### 2025-05-08 paper/paper.qmd
adding kyubins stuff

```{r}
sleep_test <- sleep[-264,]
```

>>>>

We removed variables `age` and `quality_of_sleep` and ran a MLR model. Our results showed that the reduced VIF model was very similar to the full model. Adjusted R^2 was essentially unchanged. Therefore, we concluded that `age` and `quality_of_sleep` did not contribute much to the regression analysis because they were linear combinations of other independent variables.

```{r}
sleep_test <- sleep[-264,]

sleep_test <- sleep_test %>%
  dplyr::select(!c(age, quality_of_sleep))

test_model<-lm(sleep_duration ~., sleep_test)
test_model_summary <- summary(test_model)

test_model_stats <- tibble::tibble(`F-statistic` = round(test_model_summary$fstatistic[1], 2),
                              `Adjusted R²` = round(test_model_summary$adj.r.squared, 3),
                              `Residual Sum of Squares` = round(sum(residuals(test_model)^2), 2),
                              `Total Sum of Squares` = round(sum((sleep_test$sleep_duration-mean(sleep_test$sleep_duration))^2), 2))

knitr::kable(test_model_stats, caption = "MLR Outcome for")
```

In order to avoid overfitting issues, we utilized a greedy algorithmn process to select variables that provided the lowest AIC value. We ran 3 model selection process as discussed in class and recorded the variables selected for each method. We found that all 3 methods took all of our variables in our data set. We knew that AIC penalizes the fitted model for each additional variables taken. Despite this, we saw that all variables were significant. Note that the variable selection process was done after our high VIF variables were removed.

```{r include=FALSE}
# 1. Model Selection using full_model_vif
model_backward <- step(test_model, direction = "backward")

# Define full model formula and null model for forward/stepwise
full_scope <- formula(test_model)
null_model <- lm(sleep_duration ~ 1, data = sleep_test)

model_forward <- step(null_model, scope = full_scope, direction = "forward")
model_stepwise <- step(null_model, scope = full_scope, direction = "both")

# 2. Extract variable names selected by each model
vars_backward <- attr(terms(model_backward), "term.labels")
vars_forward <- attr(terms(model_forward), "term.labels")
vars_stepwise <- attr(terms(model_stepwise), "term.labels")

# 3. Combine all variables that appeared in any model
all_vars <- sort(unique(c(vars_backward, vars_forward, vars_stepwise)))

# 4. Create a comparison table
selection_table <- data.frame(
  Variable = all_vars,
  Backward = ifelse(all_vars %in% vars_backward, "Yes", "No"),
  Forward  = ifelse(all_vars %in% vars_forward, "Yes", "No"),
  Stepwise = ifelse(all_vars %in% vars_stepwise, "Yes", "No")
)
```

```{r}
kable(selection_table, caption = "Variable Selection")
```
